---
title: 'Reproducible Research: Peer Assessment 1'
output:
  html_document: yes
    keep_md: yes
    smart: no
---


## Loading and preprocessing the data

```{r, echo = TRUE}
    
setwd("~/GitHub/RepData_PeerAssessment1")    
activity = read.csv("activity.csv", header=T)   
library("dplyr", lib.loc="~/R/win-library/3.1")    
# Create new data set with total number of steps taken per day    
summary = activity %>% group_by(date) %>% summarize(total=sum(steps, na.rm=TRUE))
str(summary)
```

## What is mean total number of steps taken per day?

```{r, echo = TRUE}   
# Create histogram for total number of steps taken per day
hist(summary$total)
```

####mean total number of steps taken per day

```{r, echo = TRUE}

mean(summary$total)

```


####median total number of steps taken per day

```{r, echo = TRUE}

median(summary$total)

```


## What is the average daily activity pattern?

```{r,echo = TRUE} 
# Create a new data set with mean steps per interval averaged accross all days
summary2 = activity %>% group_by(interval) %>%
            summarize(avg_int = mean(steps, na.rm=TRUE))
str(summary2)

plot(summary2$interval,summary2$avg_int, type="l")
```

#### Interval with maximum total steps

```{r,echo = TRUE} 
ind = which(summary2$avg_int==max(summary2$avg_int))
summary2$interval[ind]
```


## Imputing missing values
#### Calculate number of missing values

```{r,echo = TRUE}
sum(!complete.cases(activity))
```
### Imputing missing values strategy

Missing values for each interval in the original data set will be replaced with the relevant average value for that interval calculated from the original data and stored in data set "summary2".

```{r, echo = TRUE}
# Imputing missing values
for (i in 1:length(activity$steps))

        {if (is.na(activity$steps[i])) activity$steps[i] = 
                  summary2[(summary2$interval==activity$interval[i]),2]}
# Convert 'list' to 'vector'        
activity$steps = unlist(activity$steps)

# Create a new dataset with total steps per day from imputed data
total_steps_imp = activity %>% group_by(date) %>% 
                  summarize(total=sum(steps, na.rm=TRUE))

hist(total_steps_imp$total)

mean(total_steps_imp$total)

median(total_steps_imp$total)

```
### Comment
As expected the mean and median from the inputed data are larger than 

the relevant values from original data with missing values


## Are there differences in activity patterns between weekdays and weekends?
```{r}        
activity$date = as.Date(activity$date)
# Create  character vector 'weekdays' 
activity$weekdays = weekdays( activity$date)
str( activity)
# Create logical vector 'weekend'
activity$weekend = as.logical((activity$weekdays=="Saturday")|(activity$weekdays == "Sunday"))
str( activity)
# Convert logical vector 'weekend' to class 'factor'
activity$weekend = as.factor(activity$weekend)
str( activity)
# Assign factor levels' names
levels(activity$weekend) =c("weekday","weekend")

# Create a new data set with data grouped by "weekday" and "weekend"
grouped_int_day = activity %>% group_by(interval,weekend) %>% 
                  summarize(avg_steps=mean(steps))

library("ggplot2", lib.loc="~/R/win-library/3.1")
g = ggplot(grouped_int_day, aes(interval,avg_steps))
g = g + geom_line() +facet_grid(.~weekend)
g = g+xlab("Interval")+ylab("Average number of steps")   ## horizontal grid  
print(g)

g = g + geom_line() +facet_grid(weekend ~ .)   ## vertical grid
print(g)

```

### Activity patterns differences between weekdays and weekends

From the plots above it can be seen that for weekdays there is
a higher maximum for the average number of steps around the
interval '830' while for the rest of the day a higher average 
activity takes place during the weekends. 


